Forward Modeling Galaxy Size in a Cosmological Context

### 2017-10-03 paper/galsize_paper.tex
Added group-finder references

In \citet{kravtsov13}, it was found that if a stellar-to-halo mass relation is inverted to map halo mass estimates $\mhalo$ onto SDSS galaxies, and then the $\mhalo-\rvir$ relation is applied to map values of $\rvir$ onto the galaxies, then the resulting $\rhalf-\rvir$ relation of SDSS galaxies exhibits the following linear scaling across a wide range of stellar mass:
\beq
\label{eq:fiducial_model}
\rhalf = 0.0125\rvir
\eeq

Motivated by the simplicity of this scaling relation, we transform the \citet{kravtsov13} into a forward model using {\tt Halotools}. For the virial radius of halos and subhalos, we use $\rmpeak,$ the value of $\rvir$ in physical units of $\kpc$ measured at the time of peak subhalo mass, defined by

\beq
\mpeak\equiv\frac{4\pi}{3}\rmpeak^{3}\Delta_{\rm vir}(\zpeak)\rho_{\rm m}(\zpeak), 
\eeq
where for $\Delta_{\rm vir}(\zpeak)$ we use the fitting function to the ``virial" definition used in \citet{bryan_norman98}. For the model we refer to as the ``$\rvir-$only model", we add uncorrelated log-normal scatter of $\sigmarhalf=0.2$ dex to generate a Monto Carlo realization of the model population.

@@

\subsection{Progression from Backwards to Forwards Modeling}
\label{subsec:forwardsmodeling}

>>>>

Motivated by \citet{kravtsov13}, we explore a model in which $\rhalf$ is linearly proportional to halo virial radius:
\beq
\label{eq:fiducial_model}
\rhalf = 0.0125\rvir
\eeq
For the virial radius of halos and subhalos, we use $\rmpeak,$ the value of $\rvir$ in physical units of $\kpc$ measured at the time of peak subhalo mass, defined by 
\beq
\mpeak\equiv\frac{4\pi}{3}\rmpeak^{3}\Delta_{\rm vir}(\zpeak)\rho_{\rm m}(\zpeak), 
\eeq
where for $\Delta_{\rm vir}(\zpeak)$ we use the fitting function to the ``virial" definition used in \citet{bryan_norman98}. For the model we refer to as the ``$\rvir-$only model", we add uncorrelated log-normal scatter of $\sigmarhalf=0.2$ dex to generate a Monto Carlo realization of the model population.

@@

\subsection{Progression from Backward to Forward Modeling}
\label{subsec:forwardsmodeling}

Our results give an archetypal demonstration of the natural scientific progression from backward to forward modeling. In backward modeling, some mapping is applied to observed galaxies to estimate the values of model quantities such as halo mass. In \citet{kravtsov13}, the model quantities mapped onto galaxies are $\mhalo$ and $\rvir;$ other classic examples of this backward modeling include applying a group- cluster-finding algorithm to a galaxy sample to assign $\mhalo$ to the galaxies \citep[e.g.,][]{berlind_etal06,yang_etal05a}. Once the observed galaxies have been supplemented with model variables, then the relations of the galaxy-halo connection can be inferred, for example, by calculating quantities such as the mean stellar mass or quiescent fraction as a function of halo mass \citep[e.g.,][]{yang_etal05b,weinmann_etal06}.

One of the challenges of modeling galaxy size is covariance between the litany of galaxy properties that are well measured by SDSS. For example, in \citet{kravtsov13}, a halo mass estimate $\mhalo$ was mapped onto each observed galaxy, and then the $\mhalo-\rvir$ relation is applied to map values of $\rvir$ onto the galaxies. It was then observed that the resulting $\rhalf-\rvir$ relation of SDSS galaxies implied by this mapping exhibits approximately linear scaling across a wide range of stellar mass.

However, the simplicity of a linear relation by itself only gives indirect support for the assumption that $\mhalo,$ rather than $\mstar,$ statistically regulates galaxy size. By transforming the \citet{kravtsov13} scaling relation into a forward model using {\tt Halotools}, we are able to quantitatively explore the consequences of such modeling assumptions. As can be seen by comparing the dashed and dotted curves in Figure \ref{fig:clustering_ratio_upshot}, galaxy clustering gives a direct observational handle on whether $\mhalo$ or $\mstar$ is the more natural independent variable in the problem. Evidently, naive models based on $\mhalo$ rather than $\mstar$ naturally recover the qualitative trends seen in large-scale structure.

In forward modeling, the direction of inference is turned around: a mapping is instead applied to the model quantities such as $\mhalo,$ transforming the model into a mock observation. This enables a richer quantitative study of modeling hypotheses relative to backward modeling. The ability to disentangle coupled variables such as $\mstar$ and $\mhalo$ is one example of this advantage of forward modeling. Another example illustrated in the present work is the exploration of the role of splashback galaxies. In our forward modeling approach, this is entirely straightforward; in backward modeling, such an investigation would not even be possible without introducing additional modeling ingredients.

\subsection{Empirical Modeling $\rhalf:$ advantages and limitations}
\label{subsec:empirical_modeling}
